---
title: "Homework 8"
author: "Lindsey Cathcart"
date: "3/4/2020"
output: html_document
---

### 1. Examine the structure of the dataset iris
```{r}
library(dplyr)
data(iris)
head(iris)


```
##### 150 observations and 5 variables


### 2. Create new data frame iris1 
####  that contains only species virginica and versicolor with sepal lengths > 6 cm and sepal widths > 2.5 cm.

```{r}
iris1 <- filter(iris, Sepal.Length > 6, Sepal.Width > 2.5, Species %in% c("virginica", "versicolor"))

glimpse(iris1)

```
#### iris1 has 56 observations and 5 variables


### 3. Create new data frame iris2 from iris1
#### that contains only the columns for Species, Sepal.Length, and Sepal.Width

```{r}
iris2 <- select(iris1, Species, Sepal.Length, Sepal.Width)
glimpse(iris2)


```
#### iris2 has 56 observations and 3 variables


### 4. Create iris3 dataframe from iris2 
#### that orders the observations from largest to smallest sepal length

```{r}
iris3 <- arrange(iris2, by=desc(Sepal.Length))

head(iris3)


```



### 5. Create iris4 data frame from iris3
#### that creates a column with a sepal area (length*width) value for each observation

```{r}
iris4 <- mutate(iris3, Sepal.Area = Sepal.Length*Sepal.Width)
glimpse(iris4)

```
#### iris4 has 56 observations and 4 variables


### 6. Create iris5 dataframe
#### that calculates the average sepal length, the average sepal width, and the sample size of the entire iris4 data frame

```{r}
iris5 <- summarize(iris4, meanSepal.Length = mean(Sepal.Length), meanSepal.Width = mean(Sepal.Width), TotalNumber=n())
print(iris5)



```


### 7. Create iris6 dataframe
#### that calculates the average sepal length, average sepal width, and sample size for each species of the iris4 dataframe

```{r}
irisSpecies <- group_by(iris4, Species)
glimpse(irisSpecies)
iris6<- summarize(irisSpecies, meanSepal.Length = mean(Sepal.Length), meanSepal.Width = mean(Sepal.Width), TotalNumber=n())
print(iris6)
```



### Rework previous statements into an extended piping operation that uses iris at the input and generates iris6 as the output.


```{r}

library(tidyr)

iris6 <- iris %>% 
  filter(Sepal.Length > 6, Sepal.Width > 2.5, Species %in% c("virginica", "versicolor")) %>% 
  select(Species, Sepal.Length, Sepal.Width) %>% 
  arrange( by=desc(Sepal.Length)) %>% 
  mutate(Sepal.Area = Sepal.Length*Sepal.Width) %>%
  group_by(Species) %>%
  summarize(meanSepal.Length = mean(Sepal.Length), meanSepal.Width = mean(Sepal.Width), TotalNumber=n()) 

print(iris6)
```

















































