---
title: "Homework 5"
author: "Lindsey Cathcart"
date: "2/12/2020"
output: html_document
---

## Question 1
#### Assign the variable n_dims a single random integer between 3 and 10
```{r}
n_dims <- sample(3:10, 1)
print(n_dims)
```
#### Create a vector of consecutive integers from 1 to n_dims$^2$
```{r}
vector <- seq(from = 1, to = (n_dims)^2, by = 1)
print(vector)

```
#### Use the sample function to randomly reshuffle these elements
```{r}
sample(vector)
```
#### Create a square matrix with these elements
```{r}
m <- matrix(data = sample(vector), nrow=n_dims, ncol=n_dims) 
print(m)
# n_dims for dimension size regardless of the integer assigned in step 1
```
#### Find a function to transpose the matrix
```{r}
m2 <- t(m)
print(m2)
# rows and columns switch with the t() function
````


#### Calculate the sum and the mean of the elements in the first row and the last row
```{r}
# sum of first row
sum(m2[1,])
# mean of first row
mean(m2[1,])


# sum of last row
sum(m2[n_dims,])
# mean of last row
mean(m2[n_dims,])
```
#### Read about the `eigen()` function and use it on your matrix
```{r}
eigen(m2)

```
#### look carefully at the elements of `$values` and `$vectors`. What kind of numbers are these?
##### `$values` is a vector of eigenvalues for the matrix
##### `$vectors` is a matrix of eigenvectors for the matrix


#### Dig in with the `typeof()` function to figure out their type
```{r}
typeof(eigen(m2)$values)
typeof(eigen(m2)$vectors)

```



## Question 2: Create a list with the following named elements:

##### my_matrix, which is a 4 x 4 matrix filled with random uniform values

```{r}
my_matrix <- matrix(data=runif(16), nrow = 4, ncol = 4)
print(my_matrix)

```

##### my_logical which is a 100-element vector of TRUE or FALSE values. Do this efficiently by setting up a vector of random values and then applying an inequality to it.
```{r}
vector_1 <- runif(100)
my_logical <- vector_1 > 0.5
print(my_logical)

```


##### my_letters, which is a 26-element vector of all the lower-case letters in random order.

```{r}
letters_1 <- letters[1:26]
my_letters <- sample(letters_1)
print(my_letters)

```

##### create a new list, which has the element[2,2] from the matrix, the second element of the logical vector, and the second element of the letters vector.
```{r}

my_list <- list(my_matrix[2,2],
            my_logical[2],
            my_letters[2])
print(my_list)

```

##### use the typeof() function to confirm the underlying data types of each component in this list

```{r}

element1<- typeof(my_list[[1]])
element2<- typeof(my_list[[2]])
element3<- typeof(my_list[[3]])

```

##### combine the underlying elements from the new list into a single atomic vector with the c() function.

```{r}
new_vector <- c(element1, element2, element3)
print(new_vector)



```




##### what is the data type of this vector?
```{r}
typeof(new_vector)

```



## Question 3
##### Create a data frame with two variables (= columns) and 26 cases (= rows).

```{r}


##### Call the first variable my_unis and fill it with a random uniform value from 0 to 10

my_unis <- sample(seq(from = 0, to = 10, length.out = 26))
print(my_unis)

##### Call the second variable my_letters and fill it with 26 capital letters in random order.

my_letters <- sample(LETTERS[1:26])
print(my_letters)

##### Create a dataframe with these two variables
my_dataframe <- data.frame(my_unis, my_letters)
print(my_dataframe)


##### For the first variable, use a single line of code in R to select 4 random rows and replace the numerical values in those rows with NA.
my_dataframe[sample(26, 4),1] <- NA
print(my_dataframe)


##### For the first variable, write a single line of R code to identify which rows have the missing values.

is.na(my_dataframe)


###### For the second variable, sort it in alphabetical order
my_dataframe[order(my_dataframe$my_letters),]


##### Calculate the column mean for the first variable.

mean((my_dataframe$my_unis))

     
     
```



















