---
title: "Homework 7"
author: "Lindsey Cathcart"
date: "2/26/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Creating Fake Data Sets to Explore Hypotheses
### Part 1

##### Hypothesis: Decreased ground arthropod diversity is associated with invasive plant-herbivore pair.

##### Treatment 1: No established invasives (Shannon's Diversity index)
###### n = 25
###### mean = 4
###### variance = 0.22

##### Generate random normal data, visualize in histogram
```{r}
library(ggplot2)
T1Norm <- rnorm(n=25,mean=4,sd=0.22)
qplot(T1Norm, color=I("black"),fill=I("cornflowerblue"))
print(T1Norm)

```

##### Treatment 2: Established plant-herbivore pair (Shannon's Diversity index)
###### n = 25
###### mean = 1.7
###### variance = 0.35

##### Generate random normal data, visualize in histogram
```{r}
T2Norm <- rnorm(n=25,mean=1.7,sd=0.35)
qplot(T2Norm, color=I("black"),fill=I("maroon"))

```
##### Organize data into dataframe

```{r}
diversity_data <- data.frame(T1Norm,T2Norm)
print(diversity_data)

##### use tidyr gather() to make wide data long
library(tidyr)
long_diversity_df <- gather(diversity_data, Treatment, Diversity, T1Norm:T2Norm)
print(long_diversity_df)
```

##### Run ANOVA test on long_diversity_df
```{r}
anova_diversity <- aov(Diversity~Treatment, data=long_diversity_df)


summary(anova_diversity)
```


##### Generate a boxplot to visualize differences in diversity across treatment groups

```{r}
str(long_diversity_df)

Diversity <-long_diversity_df$Diversity
Treatment <- long_diversity_df$Treatment
diversity_boxplot <- boxplot(Diversity~Treatment,data=long_diversity_df,
                             xlab="Treatment",
                             ylab="Shannon's Diversity Index",
                             col=c("cornflowerblue","maroon"),
                             border="black")


```

### Part 2 
##### Adjust means to see how big the differences between groups can be to detect significance.
###### T1 mean decrease from 4 to 2.9
###### T2 mean increase from 1.7 to 2.6

##### Generate random normal data, visualize in histogram
```{r}
T1Norm <- rnorm(n=25,mean=2.9,sd=0.22)
qplot(T1Norm, color=I("black"),fill=I("cornflowerblue"))
print(T1Norm)

```

##### Treatment 2: Established plant-herbivore pair (Shannon's Diversity index)
###### n = 200
###### mean = 1.7
###### variance = 0.35

##### Generate random normal data, visualize in histogram
```{r}
T2Norm <- rnorm(n=25,mean=2.6,sd=0.35)
qplot(T2Norm, color=I("black"),fill=I("maroon"))

```

##### Organize data into dataframe
```{r}
diversity_data <- data.frame(T1Norm,T2Norm)
print(diversity_data)

##### use tidyr gather() to make wide data long
long_diversity_df <- gather(diversity_data, Treatment, Diversity, T1Norm:T2Norm)
print(long_diversity_df)
```

##### Run ANOVA test on long_diversity_df
```{r}
anova_diversity <- aov(Diversity~Treatment, data=long_diversity_df)


summary(anova_diversity)
```

##### Generate a boxplot to visualize differences in diversity across treatment groups

```{r}
str(long_diversity_df)

Diversity <-long_diversity_df$Diversity
Treatment <- long_diversity_df$Treatment
diversity_boxplot <- boxplot(Diversity~Treatment,data=long_diversity_df,
                             xlab="Treatment",
                             ylab="Shannon's Diversity Index",
                             col=c("cornflowerblue","maroon"),
                             border="black")

```

#### Part 2 result - without changing variance or sample size, means 2.6 & 2.9 are significantly different (effect size = 0.3)




### Part 3 
##### Determine the minimum sample size under which original effect size is significant.

##### Generate random normal data, visualize in histogram
```{r}
library(ggplot2)
T1Norm <- rnorm(n=2,mean=4,sd=0.22)
qplot(T1Norm, color=I("black"),fill=I("cornflowerblue"))
print(T1Norm)


T2Norm <- rnorm(n=2,mean=1.7,sd=0.35)
qplot(T2Norm, color=I("black"),fill=I("maroon"))

```

##### Organize data into dataframe

```{r}
diversity_data <- data.frame(T1Norm,T2Norm)
print(diversity_data)

##### use tidyr gather() to make wide data long
library(tidyr)
long_diversity_df <- gather(diversity_data, Treatment, Diversity, T1Norm:T2Norm)
print(long_diversity_df)
```

##### Run ANOVA test on long_diversity_df
```{r}
anova_diversity <- aov(Diversity~Treatment, data=long_diversity_df)


summary(anova_diversity)
```


##### Generate a boxplot to visualize differences in diversity across treatment groups

```{r}
str(long_diversity_df)

Diversity <-long_diversity_df$Diversity
Treatment <- long_diversity_df$Treatment
diversity_boxplot <- boxplot(Diversity~Treatment,data=long_diversity_df,
                             xlab="Treatment",
                             ylab="Shannon's Diversity Index",
                             col=c("cornflowerblue","maroon"),
                             border="black")


```

#### Minimum sample size for significance is n=2 because the difference in means is large and SD is small.



















