---
title: "homework11"
author: "Lindsey Cathcart"
date: "4/8/2020"
output: html_document
---

## Homework 11
##### 1. Repeat the exercise from the batch processing lecture using real data sets rather than purely simulated. Try to set it up as a batch process that will work on multiple file and save summary results to a common file.


##### Simulate other data sets by following the methods in Homework 6 by selecting appropriate statistical distributions and estimating parameters for those from the real data.

```{r}
library(tidyr)
library(MASS)
  
  
# clean up my dataset using gather()
mydata <- read.table("LCmmp_csv.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)

z <- gather(mydata, Year, Density, y1992:y2018)
z <- z[z$Density>0,]


str(z)
z

summary(z)

expoPars <- fitdistr(z$Density,"exponential")
rateML <- expoPars$estimate["rate"]

# data set follows the exponential distribution
# simulate new data
newdata <-data.frame( Density= rexp(243, rate = expoPars$estimate["rate"]))

head(newdata)

```


```{r}

#  -----------------------------------
# FUNCTION file_builder 
# description: creates a set of random files based off of my exponential distribution for regression  
# inputs: file_n=number of files to create
#         file_folder=name of folder for random files
#         file_size=c(min,max) number of rows
#         file_NA=average of number of NA values per column
# outputs: creates a set of random files
######################################
file_builder <- function(file_n=10,
                         file_folder="SimulatedFiles/",
                         file_size=c(100,250),
                         file_NA=0) {
for (i in seq_len(file_n)) {
  file_length <- sample(file_size[1]:file_size[2],size=1)
  
  
  
  var_x <- runif(file_length) # create random x ################## throw in exponential parameters
  var_y <- runif(file_length) # create random y
  
  
  df <- data.frame(var_x,var_y) # bind into a dataframe
  bad_vals <- rpois(n=1,lambda=file_NA) # determine NA number
  df[sample(nrow(df),size=bad_vals),1] <- NA
  df[sample(nrow(df),size=bad_vals),2] <- NA
  
  # create label for file name with padded zeroes so they are listed in numerical order
  file_label <- paste(file_folder,"ranFile", 
                      formatC(i,
                              width=3,
                              format="d",
                              flag="0"),
                      ".csv",
                      sep="")
  # set up data file and incorporate time stamp and minimal metadata
  write.table(cat("# Simulated random data file for batch processing", "\n", 
                  "# timestamp: ",as.character(Sys.time()), "\n", 
                  "#LMC", "\n", 
                  "#------------------------", "\n", 
                  "\n", 
                  file=file_label, 
                  row.names="", 
                  col.names="", 
                  sep=""))
  # now add the data frame
  write.table(x=df, 
              file=file_label,
              sep=",",
              row.names=FALSE,
              append=TRUE)
}

} # end of file_builder
#-------------------------------------


#  -----------------------------------
# FUNCTION reg_stats
# description: fits linear models, extract model stats
# inputs: 2-column data frame (x and y)
# outputs: slope, p-value, r2
######################################
reg_stats <- function(d=NULL) {
  if(is.null(d)){
    x_var <- runif(10)
    y_var <- runif(10)
    d <- data.frame(x_var,y_var)
  }
  
  . <- lm(data=d,d[,2]~d[,1])
  . <- summary(.)
  stats_list <- list(Slope=.$coefficients[2,1],
                     pVal = .$coefficients[2,4],
                     r2=.$r.squared)
  

return(stats_list)

} # end of reg_stats
#-------------------------------------



```
