---
title: "Homework 6"
author: "Lindsey Cathcart"
date: "2/19/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Homework 6

##### Generate fake data

```{r}

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

```

##### Plot a histogram of the data using ggplot 

```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

```

##### Fit a normal distribution to your data in red by first pulling out the mean and variance

###### The normal distribution doesn't fit well because the dataset has no negative values

```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute


meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat


```

##### Plot exponential probability density in blue

```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2



```
##### Plot uniform probability density using the minimum and maximum of the data in dark green

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))

 p1 + stat + stat2 + stat3


```

##### Plot gamma probabilitiy density in brown

```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4


```

##### Plot beta probability density
###### raw data will be rescaled from 0-1 and then compared to the beta distribution

```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial


```

### Part 2
##### Add real data
###### install tidyr package

```{r}


library(tidyr)


mydata <- read.table("LCmmp_csv.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
print(mydata)

z <- gather(mydata, Year, Density, y1992:y2018)
z <- z[z$Density>0,]
print(z)


str(z)
summary(z)

```

##### Plot a histogram of butterfly census 1992-2018
```{r}
p1 <- ggplot(data=z, aes(x=Density, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)


```

##### Add the empirical density curve to the histogram
```{r}

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)



```


##### Normal probability density

```{r}
normPars <- fitdistr(z$Density,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute


meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$Density),len=length(z$Density))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$Density), args = list(mean = meanML, sd = sdML))
 p1 + stat

```
##### Exponential probability density
```{r}
expoPars <- fitdistr(z$Density,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Density), args = list(rate=rateML))
 p1 + stat + stat2


```
##### Plot uniform probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="turquoise", n = length(z$Density), args = list(min=min(z$Density), max=max(z$Density)))
 p1 + stat + stat2 + stat3

```

##### Plot gamma probability density

```{r}
gammaPars <- fitdistr(z$Density,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="Orange", n = length(z$Density), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```

##### Plot beta probability density

```{r}
pSpecial <- ggplot(data=z, aes(x=Density/(max(Density + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$Density/max(z$Density + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$Density), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```
#### The exponential density curve fits my data best.

##### Get maximum likelihood parameters to simulate a new data set and plot a histogram.




```{r}

###### generate new data using parameters from exponential distribution

newdata <-data.frame( Density= rexp(243, rate = expoPars$estimate["rate"]))

head(newdata)
##### Create histogram of simulated data under gamma distribution curve
p1 <- ggplot(data=newdata, aes(x=Density, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)


expoPars <- fitdistr(z$Density,"exponential")
rateML <- expoPars$estimate["rate"]

xval <- seq(0,max(newdata$Density),len=length(z$myVar))

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$Density), args = list(rate=rateML))
 p1  + stat2
 
```

##### histogram of original data under gamma distribution curve

```{r}

p1 <- ggplot(data=z, aes(x=Density, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)








```








