---
title: "Homework 10"
author: "Lindsey Cathcart"
date: "4/1/2020"
output: html_document
---
## For loops and randomization tests

##### 1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the matrix. Finally, use return(counter) for the output.
```{r}

my_vec <- c(5,5,2,7,2,8,9,4,0,3,8,0,2,9,0,3,2,0,0,1,3,1,0,7,0,6,2,7,4,0,1)

# FUNCTION: ZeroCounter
# description: count zeros in vector
# input: my_vec = numeric vector of 31 elements with 8 zeroes
#        counter = count of "zero" elements
# output: count of zeroes
#-----------------------------------
ZeroCounter <- function(vector=my_vec) {
  counter <- 0
  for (i in seq_along(vector)){
    if(vector[i] == 0) counter = counter+1
  }
return(counter)
}

ZeroCounter(vector=my_vec)

```


##### 2. Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}

my_vec <- c(5,5,2,7,2,8,9,4,0,3,8,0,2,9,0,3,2,0,0,1,3,1,0,7,0,6,2,7,4,0,1)

# FUNCTION: ZeroCounter2
# description: count zeros in vector
# input: my_vec = numeric vector of 31 elements and 8 zeroes
# output: length of zeroes in subset of my_vec
#-----------------------------------
ZeroCounter2 <- function(vector=my_vec) {
  my_zero <- vector[vector == 0]
  
return(length(my_zero))
}

ZeroCounter2(vector=my_vec)

```


##### 3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}
r <- 2
c <- 3

##################################
# FUNCTION: DesignMatrix
# create a matrix and fill in elements 
# input: r = number of rows
#        c = number of columns
# output: matrix
#-----------------------------------
DesignMatrix <- function(r=2,
                         c=3) {
my_mat <- matrix(nrow = r, ncol = c)
for (i in 1:nrow(my_mat)){
  for (j in 1:ncol(my_mat)){
    my_mat[i,j] <- i*j
  }
}

return(print(my_mat))
}

DesignMatrix(r=2,
             c=3)

```













